Vagueness, Noise, and Signaling

### 2015-09-23 paper/version_04/paper.tex
Added a couple of diagrams to illustrate a round of play in normal sim-max games versus when imprecision is added.

@@

A sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ with much
fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
$\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity metric\footnote{A
  metric is a function of distance between any two points in a given space. It should satisfy
  certain axioms that ensure behavior that one would intuitively expect from the words
  ``similarity'' and ``distance'' alone, but these details do not matter for the purposes of
  this paper.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
and $\mystate{2}$, and a utility function $\Utils \mycolon \States \times \States \rightarrow
\mathds{R}$ such that $\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver
for a play with actual state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We
identify the receiver's acts with the states of the world, so that the game is one of guessing
the actual state, so to speak. The utility function should be a monotonically decreasing
function of similarity.

@@

If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
intends to select interpretation $\mystate{i}$, let $P_r(\mystate{r} \probbar \mystate{i})$ be
the probability with which state $\mystate{r}$ is realized.  Section~\ref{sec:exploring-rdd}
gives a concrete (parameterized) specification of $P_o$ and $P_r$ in terms of similarity of
states. For the time being, it is best to treat $P_o$ and $P_r$ as arbitrary conditional
confusion probabilities.

A single round of play of a sim-max game is then governed by five pieces of stochastic
information, where previously there were only three. At the outset, the actual state is chosen
with probability $P_a(\mystate{a})$. A randomly sampled sender observes $\mystate{o}$ with
probability $P_o(\mystate{o} \probbar \mystate{a})$ and so we expect $\messg$ with probability
$\Sstrat(\messg \probbar \mystate{o})$. (Senders respond to observations, not the actual
state!) As perception of messages is flawless, we expect a randomly sampled receiver to intend
to realize interpretation $\mystate{i}$ with probability
$\Rstrat(\mystate{i} \probbar \messg)$.  The actually realized interpretation will be
$\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$.

>>>>

  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]

  \node (1) {};
  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
  \node (3) [right=of 2] {$\messg = \Spure(\mystate{1}) \in \Messgs$};
  \node (4) [right=of 3] {$\mystate{2} = \Rpure(\messg) \in \States$};

  \path
    (1) edge node {N} (2)
    (2) edge node {S} (3)
    (3) edge node {R} (4);

  \caption{A round of play in a sim-max game: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg = \Spure(\mystate{1})$ determined by his pure strategy; a receiver (R) selects a state $\mystate{2} = \Rpure(\messg)$ determined by her pure strategy. Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
  \label{fig:sim-max-illustration}
\end{figure}

@@

Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ typically with much
fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
$\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity metric\footnote{A
  metric is a function of distance between any two points in a given space. It should satisfy
  certain axioms that ensure behavior that one would intuitively expect from the words
  ``similarity'' and ``distance'' alone, but these details do not matter for the purposes of
  this paper.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
and $\mystate{2}$, and a utility function $\Utils \mycolon \States \times \States \rightarrow
\mathds{R}$ such that $\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver
for a play with actual state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We
identify the receiver's acts with the states of the world, so that the game is one of guessing
the actual state, so to speak. The utility function should be a monotonically decreasing
function of similarity.
In its original formulation, a sender's behavior is represented by a pure strategy $\Spure \in \Messgs^\States$ which associates states to messages (given a state $\state \in \States$ which message $\messg \in \Messgs$ would be sent), and a receiver's behavior by a pure strategy $\Rpure \in \States^\Messgs$ which does the opposite.
A play in such a game is illustrated in Figure~\ref{fig:sim-max-illustration}.

@@

If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
intends to select interpretation $\mystate{i}$, let $P_r(\mystate{r} \probbar \mystate{i})$ be
the probability with which state $\mystate{r}$ is realized.  Section~\ref{sec:exploring-rdd}
gives a concrete (parameterized) specification of $P_o$ and $P_r$ in terms of similarity of
states. For the time being, it is best to treat $P_o$ and $P_r$ as arbitrary conditional
confusion probabilities.
A single round of play of a sim-max game is then governed by five pieces of stochastic
information, where previously there were only three.
An illustration is given in Figure~\ref{fig:imprecise-sim-max-illustration}.
Note that, and this is crucial, senders respond to observations not actual states, receiver strategies determine intentions not realized states, but payoff is calculated based on actual and realized states.

  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]

  \node (1) {};
  \node (2) [right=of 1,label=below:$P_a(\mystate{a})$] {$\mystate{a} \in \States$};
  \node (2') [right=of 2,label=below:$P_o(\mystate{o} \probbar \mystate{a})$] {$\mystate{o} \in \States$};
  \node (3) [right=of 2',label=below:$\Sstrat(\messg \probbar \mystate{o})$] {$\messg \in \Messgs$};
  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{i} \probbar \messg)$] {$\mystate{i} \in \States$};
  \node (4') [right=of 4,label=below:$P_r(\mystate{r} \probbar \mystate{i})$] {$\mystate{r} \in \States$};

  \path
    (1) edge node {N} (2)
    (2) edge node {} (2')
    (2') edge node {S} (3)
    (3) edge node {R} (4)
    (4) edge node {} (4');

  \caption{A round of play in a sim-max game with imprecise imitation: Nature (N) chooses an actual state with probability $P_a(\mystate{a})$; a randomly sampled sender (S) observes $\mystate{o}$ with probability $P_o(\mystate{o} \probbar \mystate{a})$ and subsequently selects $\messg$ with probability $\Sstrat(\messg \probbar \mystate{o})$; a randomly sampled receiver (R) intends to realize interpretation $\mystate{i}$ with probability $\Rstrat(\mystate{i} \probbar \messg)$ but actually realizes interpretation $\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$. Payoff for both sender and receiver is given by $\Utils(\mystate{a}, \messg, \mystate{r})$.}
  %which is typically proportional to state similarity $\similarity(\mystate{a}, \mystate{r})$.}
  \label{fig:imprecise-sim-max-illustration}
\end{figure}
